---
title: "RNA-Seq - Snail Project"
output: html_document
---

```{r}
library(dplyr)
library(pheatmap)
library(sleuth)
library(ggrepel)
```

#Differential Expression with Sleuth

```{r}
sample_id <- dir("~/Projects/20180430_rna_seq/snail_project/Kallisto")
sample_id
```

```{r}
kal_dirs <- file.path("~/Projects/20180430_rna_seq/snail_project/Kallisto", sample_id)
kal_dirs
```

```{r}
s2c <- data.frame(
  sample = sample_id,
  condition = c("GFP", "GFP", "GFP",
                "Snail", "Snail", "Snail"),
  path = kal_dirs,
  stringsAsFactors=FALSE
)
s2c
```

###Getting transcript ID-gene ID table
```{r}
mart <- biomaRt::useMart(biomart = "ENSEMBL_MART_ENSEMBL",
  dataset = "mmusculus_gene_ensembl",
  host = 'ensembl.org')
t2g <- biomaRt::getBM(attributes = c("ensembl_transcript_id", "ensembl_gene_id",
    "external_gene_name"), mart = mart)
t2g <- dplyr::rename(t2g, target_id = ensembl_transcript_id,
  ens_gene = ensembl_gene_id, ext_gene = external_gene_name)
```

```{r}
so <- sleuth_prep(s2c, target_mapping = t2g,
                  extra_bootstrap_summary=T, 
                  aggregation_column="ens_gene",
                  gene_mode=T)
```

```{r}
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, ~1, 'reduced')
```

```{r}
so <- sleuth_wt(so, which_beta="conditionSnail", which_model="full")
```

```{r}
plot_group_density(so, use_filtered = TRUE, units = "est_counts",
  trans = "log", grouping = setdiff(colnames(so$sample_to_covariates),
  "sample"), offset = 1)
```

```{r}
sleuth_table <- sleuth_results(so, 'conditionSnail', 'wt', show_all = FALSE)
```

```{r}
sleuth_significant <- dplyr::filter(sleuth_table, qval <= 0.05)
```

```{r}
write.csv(sleuth_table, file="~/Projects/20180430_rna_seq/snail_project/output/de_results.csv")
```

```{r}
genes <- unique(t2g[,2:3])
dat <- so$obs_norm
dat$scaled_reads_per_base <- NULL
dat <- spread(dat, sample, tpm)
dat <- left_join(genes, dat, by=c("ens_gene"="target_id"))
dat <- na.omit(dat)
write.csv(dat, file="~/Projects/20180430_rna_seq/snail_project/output/snail_tpm_expression.csv", row.names=F)
```

# Enrichment analysis

```{r}
library(gProfileR)
up_genes <- filter(sleuth_table, qval <= 0.05 & b >= 0.25)
down_genes <- filter(sleuth_table, qval <= 0.05 & b <= (-0.25))
```

```{r}
up_go_all <- gprofiler(query=as.character(up_genes$target_id), organism="mmusculus",
                   significant=F, src_filter=c("GO:BP", "KEGG"), hier_filtering="none")
down_go_all <- gprofiler(query=as.character(down_genes$target_id), organism="mmusculus",
                   significant=F, src_filter=c("GO:BP", "KEGG"), hier_filtering="none")

up_go <- gprofiler(query=as.character(up_genes$target_id), organism="mmusculus",
                   significant=T, src_filter=c("GO:BP", "KEGG"), hier_filtering="moderate")
down_go <- gprofiler(query=as.character(down_genes$target_id), organism="mmusculus",
                   significant=T, src_filter=c("GO:BP", "KEGG"), hier_filtering="moderate")
```

```{r}
up_go_all_sig <- gprofiler(query=as.character(up_genes$target_id), organism="mmusculus",
                   significant=T, src_filter=c("GO:BP", "KEGG"), hier_filtering="none")
down_go_all_sig <- gprofiler(query=as.character(down_genes$target_id), organism="mmusculus",
                   significant=T, src_filter=c("GO:BP", "KEGG"), hier_filtering="none")

write.csv(up_go_all_sig, file="~/Projects/20180430_rna_seq/snail_project/output/go_up_all_sig_terms.csv")
write.csv(down_go_all_sig, file="~/Projects/20180430_rna_seq/snail_project/output/go_down_all_sig_terms.csv")
```

## Plotting
I want to do a dot plot thing, where the size/colour of the dot is the z-score used by GOPlot (hits A).
```{r}
#Z-score = (Up hits) - (Down hits) / sqrt(term size)
go_terms <- inner_join(up_go_all, down_go_all, by="term.id") #Note that we lose a couple IDs
go_terms_z <- data.frame(term.id = go_terms$term.id,
                         term.name = go_terms$term.name.x,
                         domain = go_terms$domain.x,
                         term.size = go_terms$term.size.x,
                         up_overlap = go_terms$overlap.size.x,
                         down_overlap = go_terms$overlap.size.y,
                         pval_up = go_terms$p.value.x,
                         pval_down = go_terms$p.value.y,
                         z_score = (go_terms$overlap.size.x - go_terms$overlap.size.y) / sqrt(go_terms$term.size.x))
#This is a master table
```

Now to get the hierarchically-filtered terms
```{r}
go_terms_sig <- filter(go_terms_z, term.id %in% up_go$term.id |
                         term.id %in% down_go$term.id)
#Prettying it up and ordering
go_terms_sig$term.name <- Hmisc::capitalize(as.character(go_terms_sig$term.name))
go_terms_sig$term.name <- factor(go_terms_sig$term.name, 
                                 levels=go_terms_sig$term.name[order(go_terms_sig$z_score)])
```

Plot
```{r}
bluered <- rev(colorRampPalette(brewer.pal(9, "RdBu"))(100))
go_plot <- ggplot(go_terms_sig, aes(x=z_score, y=term.name)) +
  geom_point(size=2, aes(color=z_score)) +
  geom_vline(xintercept=0, linetype=2) +
  scale_colour_gradientn(colours = bluered, name="Z-Score",
                         limits=c(-1*(max(go_terms_sig$z_score)), max(go_terms_sig$z_score))) +
  ylab("") + xlab("Term Z-Score") +
  theme_classic() +
  theme(legend.title=element_text(size=10),
        legend.position="none") #Not sure if I want the color bar yet
ggsave(go_plot, file="~/Projects/20180430_rna_seq/snail_project/figs/snail_go_plot.pdf",
       width=4.75, height=2.5)
ggsave(go_plot, file="~/Projects/20180430_rna_seq/snail_project/figs/snail_go_plot.png",
       width=4.5, height=2.5)
```

# Volcano plot

```{r}
#Label list
label_list <- c("Snai1", "Col6a2", "Col18a1", "Col6a1", "Col1a1", "Alcam", "Itga7", "Krt18", "Adamts4")
```


```{r}
library(ggrepel)
volcano_dat <- sleuth_table
volcano_dat$sig <- "No"
volcano_dat$sig[which(volcano_dat$qval <= 0.05 & abs(volcano_dat$b) > 0.25)] <- "Yes"

volcano_dat$labels <- ""
label_rows <- which(volcano_dat$ext_gene %in% label_list)
volcano_dat$labels[label_rows] <-volcano_dat$ext_gene[label_rows]

volcano <- ggplot(volcano_dat, aes(x=b, y=-log10(qval))) +
  geom_point(size=2, alpha=0.5, aes(color=sig)) +
  geom_vline(xintercept=c(-0.25, 0.25), linetype=2) +
  geom_hline(yintercept=-log10(0.05), linetype=2) +
  scale_x_continuous(limits=c(min(volcano_dat$b), (-1)*min(volcano_dat$b))) +
  geom_text_repel(aes(label=labels),
                    direction = "both",
                    size=4,
                    segment.alpha=0.75,
                    segment.color="grey50",
                  min.segment.length = 0.1,
                  force=0.035,
                  nudge_y=2) +
  xlab("Fold Change (Beta)") + ylab("-log10(q value)") +
  scale_color_manual(values=c("grey40", "cyan3")) +
  theme_classic() +
  theme(legend.position="none",
        axis.text=element_text(size=12, color="black"),
        axis.title=element_text(size=14, color="black"))
ggsave(volcano, file="~/Projects/20180430_rna_seq/snail_project/figs/snail_volcano_plot.pdf",
       width=4, height=3.25)
ggsave(volcano, file="~/Projects/20180430_rna_seq/snail_project/figs/snail_volcano_plot.png",
       width=4, height=3.25)
```

# GO Term plots of TGFB1-specific GO terms
I'm just pulling terms from a comparison Lauren did. She took the genes that TGFB induces but snail does not, and then did GO terms.

```{r}
bluered <- rev(colorRampPalette(RColorBrewer::brewer.pal(9, "RdBu"))(100))
go_terms <- data.frame(terms = c("Positive regulation of cell migration",
                                 "Cell adhesion",
                                 "Wound healing",
                                 "Multicellular organism development",
                                 "Cell response to TGFB stimulus",
                                 "Negative regulation of cell death",
                                 "Immune system process",
                                 "Oxidation-reduction process",
                                 "Regulation of apoptotic process",
                                 "Positive regulation of T cell activation",
                                 "Negative regulation of JAK/STAT",
                                 "Positive regulation of cell death"),
                       p.val = c(9.9e-13, 5e-11, 9e-7, 2e-6, 1.4e-5, 4e-5,
                                 1.1e-7, 5.3e-5, 8.3e-4, 8.6e-4, 8.6e-4, 1.6e-3),
                       direction = c(rep("Up", 6), rep("Down", 6)))
go_terms$terms <- factor(go_terms$terms, levels=rev(go_terms$terms))
go_plot <- ggplot(go_terms, aes(x=terms, y=-log10(p.val))) +
  geom_col(color="grey30", width=0.035) +
  geom_point(size=2, aes(color=direction)) +
  geom_hline(yintercept=-log10(0.05), linetype=2) +
  scale_colour_manual(values=c(bluered[15], bluered[85])) +
  scale_y_continuous(limits=c(0,12.5), expand=c(0,0)) +
  xlab("") + ylab("-log10(p-value)") +
  theme_classic() +
  theme(legend.position="none",
        plot.margin=unit(c(0.5,0.5,0.5,0.5), "cm")) +
  coord_flip()
ggsave(go_plot, file="~/Projects/20180430_rna_seq/snail_project/figs/tgfb_snail_independent.pdf",
       width=5, height=2.5, useDingbats=F)
ggsave(go_plot, file="~/Projects/20180430_rna_seq/snail_project/figs/tgfb_snail_independent.png",
       width=5, height=2.5)
```

